---
title: "icareview1"
author: "Reesa John"
date: "March 11, 2019"
output: html_document
---

You will want to think about how your code should handle numbers that are negative/zero, and
numbers that are larger than 26.

There is a pre-loaded vector called letters that contains a-z.
If your function is successful, the code below should return a real word.
caesar("sfzqlov", -23)
```{r}
library(tidyverse)
```

97-122 a-z
```{r}
caesar<- function(string, shift){
  shift <- shift %% 26
  stringNum <- utf8ToInt(string)
  for(i in seq_along(stringNum)){
    
    stringNum[i] = stringNum[i]+shift
    
    if(stringNum[i]>122){
      stringNum[i] = stringNum[i]-26  
    }
    else if(stringNum[i]<97){
      stringNum[i] = stringNum[i]+26
    }

  }
  intToUtf8(stringNum)
  
}

caesar("sfzqlov", -23)
```

```{r}
message <- read.csv("https://www.dropbox.com/s/ctplwmkvey4yo0e/cs_encrypted.txt?dl=1",stringsAsFactors = FALSE, header = TRUE)$x
```

Task Two: Decode a message
The encrypted message can be loaded using the code below. This message has been encrypted via the Caesar
cipher, with an unknown shift.
```{r}
for(i in 1:26){
  print(paste(caesar(message[1],i),i))
}

```
shift is 9

2. Use map or similar to decode the full message.

```{r}
temp <- map(message,caesar,shift = 9)
paste(temp,collapse = " ")
```

3. On Polylearn, submit the name of the historical figure who gave the speech in the message