---
title: "Lab2"
author: "Reesa John"
date: "1/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
avo <- read.csv("https://www.dropbox.com/s/vsc1dkosz6nwake/avocado.csv?dl=1")

## put this at the beginning of all your Rmarkdown projects:

library(tidyverse)
```
Preliminary Info:
Product Lookup Code (PLU): (X)4046 - Total # of Small Hass #60 sizes and smaller sold
                           (X)4225 - Total # of Large Hass #40 & #48 sizes sold
                           (X)4770 - Total # of X-Large Hass #36 sizes and larger sold
                           
X1 does not seem to be a meaningful variable

Exploring the dataset:
```{r}
avo <- avo %>% select(-X1)
summary(avo)
```

1) Which region sold the most bags of small organic avocados in 2017?
Hint: TotalUS does not count as a region!

The region that sold the most bags of small organic avocados in 2017 is the Northeast.

```{r}
avo %>% 
  filter(region!="TotalUS",type =="organic", year ==2017)%>%
  top_n(1,Small.Bags)%>%
  select(region)
```


2) Use separate() to split the Date variable into year, month, and day. In which month is the highest
volume of avocado sales?

On average, January tends to sell the highest volume of avocadoes, though the highest monthly avocado sales in any year was in the March of 2018.
```{r}
tempAvo1 <-  avo %>%
  select(-year)%>%
  separate(Date,into = c("year","month","day"))%>%
  group_by(month,year)%>%
  summarize(monthlyVolume=sum(Total.Volume))

tempAvo1 %>%
  group_by(month)%>%
  summarize(averageMonthlyVolume = mean(monthlyVolume))%>%
  top_n(1,averageMonthlyVolume)

tempAvo1%>%
  unite("YearMonth", c(month,year), sep = "-")%>%
  top_n(1,monthlyVolume)
  
  

```


3) Which regions sell the most avocados by volume? Plot side-by-side boxplots of Total Volume for only
the 5 regions with the highest averages for the Total Volume variable.
Hint: Once you narrow down to the top 5 regions, you can use pull() to save the vector of region names
for later use.

The five regions that sell the most avocados are California, Northeast, South Central, Southeast, and the West, though not in that order necessarily.
```{r}
top5 <- avo %>%
  filter(region!="TotalUS")%>%
  group_by(region)%>%
  summarize(averageMonthlyVolume = mean(Total.Volume))%>%
  top_n(5,averageMonthlyVolume) %>% 
  arrange(desc(averageMonthlyVolume))%>%
  pull(region)



avo %>% 
 filter(region %in% top5) %>%
 ggplot(aes(x = region, y = Total.Volume, fill=region)) +ggtitle("Volume of Avocadoes Sold By Region")+xlab("Region")+ylab("Total Volume")+labs(fill="Region")+
 geom_boxplot() 
                     
```


The following four California regions are in this dataset: LosAngeles, SanDiego, Sacramento, SanFrancisco.
Answer the following questions about the California regions only.
Hint: These questions will require restructuring of your data!

```{r}
CalRegions <- c("LosAngeles", "SanDiego","Sacramento","SanFrancisco")
CalRegions <- avo%>%
  filter(region %in% CalRegions)

```


4) In which regions is the price of organic versus conventional avocados most different? Support your
answer with a few summary numbers and a plot.

Looking at the plot and table, San Francisco has the highest difference in average prices for organic vs conventional avocados at by around an 85 cents difference.

```{r}
diffByRegion <-  CalRegions %>%
  select(AveragePrice,type,region)%>%
  group_by(region,type)%>%
  summarize(price = mean(AveragePrice))%>%
  ungroup()%>%
  spread(type,price)%>%
  mutate("organicConventional" = organic-conventional)

diffByRegion %>% group_by(region)%>%summarize(organicConventional)

diffByRegion%>%ggplot(aes(x=region,y=organicConventional,fill=region))+ggtitle("Region vs Difference in Organic and Conventional Avocado Prices")+xlab("Region")+ylab("Organic - Conventional")+labs(fill="Region")+
  geom_col()
  
```


5) How do their avocado habits differ? Make a plot that shows, for all 4 California regions, the percent
of avocado sales that are small, large, or extra large. Separate your plot by conventional vs. organic

```{r}

avoHabits <- CalRegions %>%
  group_by(type,region)%>%
  select(region,type,Small.Bags,Large.Bags,XLarge.Bags,Total.Bags)%>%
  mutate(smallPct = Small.Bags/Total.Bags, largePct = Large.Bags/Total.Bags, xLargePct=XLarge.Bags/Total.Bags)%>%
  summarize_at(vars(smallPct,largePct,xLargePct,Total.Bags), funs(mean), na.rm = TRUE)%>%
  gather(key=typePct,percent,c("smallPct","largePct","xLargePct"))
avoHabits%>%
  ggplot(aes(x=region,y=percent,fill=typePct))+geom_col()+facet_grid(~type)+ggtitle("Avocado Habits")+
  xlab("Region")+ylab("Percent")+scale_fill_discrete(name="Bag Type",
                         breaks=c("smallPct","largePct","xLargePct"),
                         labels=c("Small Bags", "Large Bags", "Extra Large Bags"))+scale_x_discrete(labels = c("Los\nAngeles","Sacramento","San\nDiego","San\nFrancisco"))

```




