---
title: "Lab3Reesa"
author: "Reesa John"
date: "2/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(lubridate)
songs <- read.table("https://www.dropbox.com/s/jrwjthqo9b5o07g/billboard_songs.txt?dl=1",header = TRUE,
stringsAsFactors = FALSE)
```

```{r}

#pattern = "y$" y at end

songs <-songs %>%
  mutate(chart.date = as.character(chart.date))%>%
  mutate(chart.date = parse_date_time(chart.date,orders = "Ymd"))%>% 
  mutate(featuring = case_when(
    str_detect(artist,pattern="featuring") ~ str_replace(artist,pattern = ".* featuring ",replacement = "")))%>%
  mutate(collaboration = case_when(
    str_detect(artist,"&") & !str_detect(artist,pattern="featuring") ~ TRUE,
    TRUE ~ FALSE
    )
  )

```

1) What 10 songs spent the longest on the charts? Give only the title, artist, and weeks.

These are the top 10 songs that have spent the longest on the charts.

```{r}
songs %>%
  arrange(desc(weeks.on.chart))%>%
  select(title,artist,weeks.on.chart)%>%
  top_n(10)
  
```

2) What date did the oldest song(s) in this dataset leave the charts? Give only the date.

The date is July 20th, 1940.
```{r}
songs %>% 
  arrange(chart.date)%>%
  top_n(wt=chart.date,-1)%>%
  select(chart.date)%>%
  unique()

```

3) What songs could have been played at your 16th birthday party? That is, which songs overall.peaked at #1 and entered the charts within a couple months (before or after) your 16th birthday? Give only
the song title, artist, and date of chart entry.

I'm looking at songs that entered the charts within 3 months of my 16th birthday, May 10th.

```{r}

near_my_birthday = ymd("2015-02-01")%--% ymd("2015-09-01")

songs %>%
  mutate(entered_chart = chart.date - weeks(weeks.on.chart))%>%
  filter(entered_chart %within% near_my_birthday)%>%
  select(title,entered_chart)

```

4) What is the most common word, at least 4 letters long, used in the title of any song? Give only the
word itself, and its count.

Love is the most common word.
```{r}
titles <- songs %>%
pull(title)

four_char_words <- titles %>%
  str_split(pattern = " ")%>%
  unlist() %>%
  str_subset(pattern = ".{4,}")
  
  
four_char_words %>%
table() %>%
sort() %>%
rev() %>%
head(1)

```

5) 
For this question, you may ignore collaborations; i.e., “Ariana Grande” is considered a different artist from
“Jessie J, Ariana Grande & Nicki Minaj”
Hint: the function n() might help you summarize.

Which five artists were the most successful in terms of:

• Number of songs that were #1 hits

The Beatles, Mariah Carey, Elvis Presley, Madonna, Michael Jackson would be.

```{r}
songs %>% 
  filter(overall.peak==1)%>%
  select(artist,overall.peak)%>%
  group_by(artist)%>%
  summarize(totalTop = sum(overall.peak))%>%
  arrange(desc(totalTop))%>%
  top_n(5)
  
```

• Total weeks any of their songs spent on the chart.

Elvis Presley, Elton John, Madonna, Perry Como, Kenny Chesney would be.

```{r}

songs %>% 
  select(artist,weeks.on.chart)%>%
  group_by(artist)%>%
  summarize(totalWeeks = sum(weeks.on.chart))%>%
  arrange(desc(totalWeeks))%>%
  top_n(5)

```


6) Which artist has featured on the most Billboard charting songs? 

Lil Wayne, who is featured in 42 songs.

```{r}
features <- songs %>%
  filter(!is.na(featuring))%>%
  pull(featuring)
  
features <- unlist(str_split(features, pattern =fixed(",")))

features <- unlist(str_split(features,pattern=fixed("&")))

as.data.frame(features)%>%
  group_by(features)%>%
  count()%>%
  arrange(desc(n))%>%
  head(1)
  

  
```

Which artist has collaborated on the most Billboard charting songs?

His Orchestra has collaborated on the most Billboard charting songs.

```{r}

collabs <- songs %>%
  filter(collaboration) %>%
  mutate(artist = str_trim(artist))%>%
  pull(artist)


collabs <- unlist(str_split(collabs, pattern =fixed(",")))
collabs <- unlist(str_split(collabs,pattern=fixed("&")))
collabs <- unlist(str_split(collabs,pattern=fixed("/")))

as.data.frame(collabs)%>%
  mutate(collabs = str_trim(collabs))%>%
  group_by(collabs)%>%
  count()%>%
  arrange(desc(n))%>%
  head(1)
```